Modelling last glacial cycle ice dynamics in the Alps

### 2018-07-20 articles/cycle/response-referee1.tex

    We did not know about references on MIS~3 and have developped our sentence
    to better reflect the current state of knowledge on palaeo-precipitation:

    \msquote{%
        Palaeoclimate proxies indicate slightly reduced LGM precipitation in
        western Europe with anomalies diminishing eastwards
        \citep{Wu.etal.2007}. Regional circulation models indicate generally
        dryer conditions during MIS~3 \citep{Barron.Pollard.2002,
        Kjellstrom.etal.2010} but more precipitation south of the Alps during
        MIS~2 LGM \citep{Strandberg.etal.2011, Ludwig.etal.2016}.}

    \referee{%
        The past climate variations are prescribed following 3 quite distal
        core records, and the most distal (EPICA) is chosen as yielding the
        best fit to Alpine glacial evidence. The basis for preferring EPICA
        seems reasonable (matching some higher-frequency amplitudes of ice
        variability, section 3.3). However, this agreement is in a sense
        coincidental, in that there is no direct meteorological link between
        Antarctic and Alpine regional climate variations. Are there any
        proximal proxy records of Alpine climate at all, perhaps lacustrian
        varves, that could be used to assess the EPICA-based shifts in air
        temperatures and precipitation, even over limited periods of the last
        120\,kyrs?}

    A short review of available proximal proxy records was added here:

    \msquote{%
        Only few regional proxy records exist that extend over periods when the
        Alps were glaciated \citep{Heiri.etal.2014}. These include lake
        sediment records in north and west of the Alps
        \citep[e.g.,]{Beaulieu.Reille.1992, Wohlfarth.etal.2008,
        Duprat-Oualid.etal.2017}, and cave speleothem in the Eastern
        \citep[e.g.,][]{Spotl.Mangini.2002, Boch.etal.2011} and Western
        \citep{Luetscher.etal.2015} Alps. Due to the scarcity of vegetation
        north of the Alps during glacial periods, varying sources for moisture
        advection, and the limited duration of the records, quantitative
        palaeoclimatic interpretation will require combining multiple proxies
        in space and time, and comparing them against regional circulation
        model output \citep{Heiri.etal.2014}.}

    In the review paper by \citet{Heiri.etal.2014}, the latter has been
    specifically identified as a neeeded future development, but to our
    knowledge, no quantitative reconstruction has been made available since
    then. The following word of caution was also added in the conclusions:

    \msquote{%
        This suprising result is likely coincidental as there is no direct link
        between European and Antarctic climate. This highlights the need for
        more quantitative reconstructions of European palaeoclimate.}

### 2018-07-20 articles/cycle/alps-cycle.tex
Assess referee 1 comment on regional proxies.

    Temperature offset time-series, ${\Delta}T_{\text{TS}}$, are derived from
    palaeo-temperature proxy records from the Greenland Ice Core Project
    \citep[GRIP;][]{Dansgaard.etal.1993}, the European Project for Ice Coring
    in Antarctica \citep[EPICA;][] {Jouzel.etal.2007}, and an oceanic sediment
    core from the Iberian margin \citep[MD01-2444;][]{Martrat.etal.2007}.
    Palaeo-temperature anomalies from the GRIP record are calculated from the
    oxygen isotope (\chem{\delta^{18}O}) measurements using a~quadratic
    equation \citep{Johnsen.etal.1995},
    %
    \begin{align}
      {\Delta}T_{\text{TS}}(t) ={~}&-11.88 [\chem{\delta^{18}O}(t)
                                    -\chem{\delta^{18}O}(0)] \nonumber \\
                                   &-0.1925 [\chem{\delta^{18}O}(t)^2
                                    -\chem{\delta^{18}O}(0)^2] \,,
    \end{align}
    %
    while temperature reconstructions from the EPICA and MD01-2444 records are
    provided as such. For each proxy record used and each of the parameter
    set-ups used in the sensitivity tests, palaeo-temperature anomalies are
    scaled linearly (Table~\ref{tab:records}, Fig.~\ref{fig:timeseries}a) so
    that, within a $150\times100$\,km rectangular domain covering the Rhine
    glacier piedmont lobe (Fig.~\ref{fig:footprints}a, black rectangle), the
    modelled cumulative glaciated area during MIS~2 (29--14\,ka) is
    consistent with the glaciated area of the geological reconstruction
    \citep{Ehlers.etal.2011}.

@@

@@

    However, these results are limited by uncertainties on ice physics. The
    the till deformation model used here does not hold for sliding over
    bedrock surfaces. On the other hand, the constant friction angle used is
    representative of wet till but weaker basal conditions may have applied
    over saturated lake sediments where they occured. A constant rheology was
    used for temperature ice containing more than 1\,\% of liquid water.

>>>>

    Only few regional proxy records exist that extend over periods when the Alps
    were glaciated \citep{Heiri.etal.2014}. These include lake sediment records
    in north and west of the Alps \citep[e.g.,]{Beaulieu.Reille.1992,
    Wohlfarth.etal.2008, Duprat-Oualid.etal.2017}, and cave speleothem in the
    Eastern \citep[e.g.,][]{Spotl.Mangini.2002, Boch.etal.2011} and Western
    \citep{Luetscher.etal.2015} Alps. Due to the scarcity of vegetation north
    of the Alps during glacial periods, varying sources for moisture advection,
    and the limited duration of the records, quantitative palaeoclimatic
    interpretation will require combining multiple proxies in space and time,
    and comparing them against regional circulation model output
    \citep{Heiri.etal.2014}.

    In our simulations, temperature offset time-series,
    ${\Delta}T_{\text{TS}}$, are derived from distal
    palaeo-temperature proxy records from the Greenland Ice Core Project
    \citep[GRIP;][]{Dansgaard.etal.1993}, the European Project for Ice Coring
    in Antarctica \citep[EPICA;][] {Jouzel.etal.2007}, and an oceanic sediment
    core from the Iberian margin \citep[MD01-2444;][]{Martrat.etal.2007}.
    Palaeo-temperature anomalies from the GRIP record are calculated from the
    oxygen isotope (\chem{\delta^{18}O}) measurements using a~quadratic
    equation \citep{Johnsen.etal.1995},
    %
    \begin{align}
      {\Delta}T_{\text{TS}}(t) ={~}&-11.88 [\chem{\delta^{18}O}(t)
                                    -\chem{\delta^{18}O}(0)] \nonumber \\
                                   &-0.1925 [\chem{\delta^{18}O}(t)^2
                                    -\chem{\delta^{18}O}(0)^2] \,,
    \end{align}
    %
    while temperature reconstructions from the EPICA and MD01-2444 records are
    provided as such. For each proxy record used and each of the parameter
    set-ups used in the sensitivity tests, palaeo-temperature anomalies are
    scaled linearly (Table~\ref{tab:records}, Fig.~\ref{fig:timeseries}a) so
    that, within a $150\times100$\,km rectangular domain covering the Rhine
    glacier piedmont lobe (Fig.~\ref{fig:footprints}a, black rectangle), the
    modelled cumulative glaciated area during MIS~2 (29--14\,ka) is
    consistent with the glaciated area of the geological reconstruction
    \citep{Ehlers.etal.2011}.

@@

    This suprising result is likely coincidental as there is no direct link
    between European and Antarctic climate. This highlights the need for more
    quantitative reconstructions of European palaeoclimate.

@@

    However, these results are limited by uncertainties on ice physics. The
    till deformation model used here does not hold for sliding over
    bedrock surfaces. On the other hand, the constant friction angle used is
    representative of wet till but weaker basal conditions may have applied
    over saturated lake sediments where they occured. A constant rheology was
    used for temperature ice containing more than 1\,\% of liquid water.
